Apparent preferences for cognitive effort fade when multiple forms of effort

### 2020-11-17 Cost_manuscript.Rmd
giving extra form to discussion (will probably change a bit once modeling results are evaluated)

  Understanding the sporadic attractiveness of effortful options remains elusive. In this study, we tested the hypothesis that such appeal could depend on the ability to achieve the same outcome through different means. Using two foraging experiments that exposed individuals to a single or intermixed types of demands, we found that 1) preferences for cognitive effort were stable when individuals face a single form of demand, but fade over time when other costs are also experienced; and 2) the phenomenon can be explained by a model in which preferences depend on how participants construe the amount of time a task takes, which could hypothetically be recalibrated when all tasks are interleaved.

  Our findings provided partial support for our pre-registered hypotheses, which can be found in https://osf.io/2rsgm/registrations. In accord with our predictions, decision makers in both experiments successfully approximated reward-maximizing strategies, as evidenced by acceptance rates near optimal points. This is in line with previous foraging experiments, where the main deviation from optimality has been a tendency to overharvest [constantino; hayden; wikenheiser; sweiss]. This successful integration of the task was further evidenced by the ability of participants to post-experimentally describe near-optimal strategies, discarding the possibility of our results being driven by noise or bias in the estimation of environmental richness [hayden; garrett; dundon]. In contrast, our prediction that effort requirements would produce more acceptances was only true for the cognitive group, and this tendency for higher acceptances did not switch to passive delay as predicted for the within-subjects experiment. These hypotheses originally stemmed from a combination of pilot data, and prevalent theorizing that incentivized work should become more appealing [Eisenberger; Inzlicht], but that people would opt for the easiest available way to achieve desired outcomes [@Hull1943; Olivola]. Instead, our novel use of foraging paradigms, in addition to our inclusion of types of effort that are often studied separately, highlighted task-specific attributes contributing to the attractive or aversive nature of demands. The fact that preferences for cognitive effort were stable as predicted in the first experiment, and present at the beginning of the second experiment, further supported this idea. However, the gradual convergence in acceptances in the second experiment emphasized that these features could be contextually molded.

  These choice dynamics ruled out a number of explanatory candidates. For instance, differences in choice stability between experiments suggest that a demand's cost is not intrinsic. Effort preferences due to boredom would not have shown the global acceptance decline and convergence in the within-subjects experiment, which also discards the possibility that all our participants were high on need for cognition (i.e. Ceyda; Laura; Cacioppo et al 1996). Even though fatigue could explain the decrease in effort acceptances across experiments, we would have expected tired participants to turn to wait trials over time in order to preserve earning rates. We also ruled out the sunk cost effects used to explain overharvesting (Wikenheiser2013; Sweiss2018), as participants expressed their choices as early as possible, resulting in occassional underharvesting. Even though the reciprocal relationship between effort and outcome can often explain preferences for effort (learned industriousness, Inzlicht2018), the observed differences in acceptances between effort types and the eventual convergence of rates in the within-subjects study cannot be explained by reward alone, as demands were rewarded at the same rate.

  We do not discard the influence of these elements in daily effort evaluations. Instead, we computationally highlight that effort can became valuable by shortening the subjective elapsed time of trials, a perception that could be recalibrated as individuals experienced multiple forms of demands. Alterations in subjective time sensitivity have been shown to play a role in temporal discounting [@Mckerchar2003], and are often experienced when we are fully immersed in a task [@czksdmsdkjdc]. The shorter time sensitivity of cognitive effort could be due to its constant recruitment of working memory, impeding estimations of the nominal time, or because of the discrete nature of the task (versus a continuous configuration of the physical and wait tasks). These results expand our understanding of the attractiveness of effort, bringing forth outcome-independent features that can be leveraged to motivate effort engagement in diverse everyday scenarios.

  The use of foraging paradigms....

  Rather than comparing discounting functions among demands, allowing people to choose for them helped us control for a number of confounds. For example, effortful tasks often take longer to perform in everyday life. We eliminated this possibility by using a common handling time for all demands.

@@

- The observed choice dynamics ruled out a number of explanatory candidates. For instance, differences in choice stability between experiments suggest that a demand's cost is not intrinsic. Effort preferences due to boredom would not have shown the global acceptance decline and convergence in the within-subjects experiment, which also discards the possibility that all our participants were high on need for cognition (i.e. Ceyda; Laura; Cacioppo et al 1996). Even though fatigue could explain the decrease in effort acceptances across experiments, we would have expected tired participants to turn to wait trials over time in order to preserve earning rates. We also ruled out the sunk cost effects used to explain overharvesting (Wikenheiser2013; Sweiss2018), as participants expressed their choices as early as possible, resulting in occassional underharvesting. Noise also discarded (Hayden paper), as values on each end of the reward range moved little over time. Even though the reciprocal relationship between effort and outcome can often explain preferences for effort (learned industriousness, Inzlicht2018), the observed differences in acceptances between effort types and the eventual convergence of rates in the within-subjects study cannot be explained by reward alone, as demands were rewarded at the same rate. Instead, we proposed that subjective time sensitivity has been shown to play a role in temporal discounting, improving on classic hyperbolic fits (Mckelachar....). Our findings do not discard the influence of these elements in daily effort evaluations. It supplements our understanding of what drives peoples preferences.
- We still think that cognitive and physical manipulations were effortful. Our physical effort task has been repeatedly used to probe effort. Similar task switching has been shown to feel cognitively effortful (Kool). Furthermore, the effortful nature of our manipulations is evidenced by the steeper decrease in acceptances over time at longer handling times (Treadway). There are many reasons why such cognitive manipulations can be effortful, which are often centered in the amount of control. Or the OC. 
- Note that the calibration of efforts were different. We adjusted to people's physical capabilities, but trained them in the cognitive tasks until they reached criterion. 
- Efficacy, along with reward amount, can influence how much control is exherted.
- But here we do not focus on choosing the *amount* of cognitive control to be exherted during performance, which can be due to efficacy, agency, or reward amount. Instead, we focus on what makes people *pursue* the task in spite of how effortful it is.

@@

- Control/Exp Design: Having a between-subjects experiment provided a baseline of foraging behavior that helped us interpret the findings of the second experiment
- Control/Exp Design: for time as confound of effort
- Control/Exp Design: having two efforts helped dissociate domain-general from task-specific features.
- Control/Exp Design: participants were fully aware. This removed the possibility of asymmetric learning and choice difficulty. Demonstrated by stable strategies.
- Control/Exp Design: in demand experiments, chosen options are often carried out after the experiment, but there are known discrepancies between expected and obtained value that could affect repeated decisions in real life. Alt from grant: Investigations of the pursuit and receipt stages have been comparatively limited. Many studies have sought to isolate the evaluation stage by presenting subjects with choices whose results do not come into effect until the end of the experimental session (with delays spanning up to months), deliberately setting aside the question of how the ongoing experience of cost and outcome affects subsequent choices [5,11,15,16,19]. However, discrepancies between expected and experienced value are well known [20], and choosing immediate outcomes could be due to factors other than time preference, such as the risk of being unable to redeem the future rewards [3,21,22]. Relying on a foraging paradigm allowed us to control for these and other variables...then control/exp design?
- Control/Exp Design: even in TAFC tasks, selecting the shorter option could index a wish to find a better set of options
- Control/Exp Design: Optimality let us know that people acted as if effort was attractive. This was evidenced by both the lower earnings by the cognitive effort group, and by the base OC model. Of note is that the optimality of the strategies produced by each cost depended on the richness of the block. This can help explain why the wait group earned the most despite having similar acceptance rates to the physical group, as the former was closer to optimality in the richest environment.
- This appeal couldn't be explained by a number of explanatory candidates
- This includes the often-cited reciprocal relationship between effort and reward
- But it turns out subjective time fits the bill. 
- Cognitive effort gained value by making trials feel shorter
- This could be because: 1) discrete nature of cog taska llowed for counting; 2) flow; 3) persistent recruitment of working memory (Buhi and Gershman). Talk about future experiments: if this is subjective time, changing the pacing of the tasks should change the acceptance rates for cognitive effort.

@@

- AFAIK, Eisenberger1982 provides evidence that task diversity motivates further effort engagement

>>>>

  Understanding the sporadic attractiveness of effortful options remains elusive. In this study, we tested the hypothesis that the perceived cost or value of a demand could depend on the ability to achieve the same outcome through different means. Using two foraging experiments that exposed individuals to a single or intermixed types of demands, we found that 1) preferences for cognitive effort were stable when individuals faced a single form of demand, but faded over time when other costs were also experienced; and 2) the phenomenon can be explained by a model in which preferences depend on how participants construe the amount of time a task takes, which could hypothetically be recalibrated when tasks are interleaved.

  Our findings provided partial support for our pre-registered hypotheses, which can be found in https://osf.io/2rsgm/registrations. In accord with our predictions, decision makers in both experiments successfully approximated reward-maximizing strategies, as evidenced by acceptance rates near optimal points. This is in line with previous foraging experiments, where the main deviation from optimality has been a tendency to overharvest [constantino; hayden; wikenheiser; sweiss]. This successful integration of the task was further evidenced by the ability of individuals to describe near-optimal strategies after participation, discarding the possibility of our results being driven by noise or bias in the estimation of environmental richness [hayden; garrett; dundon]. In contrast, our prediction that effort requirements would produce more acceptances was only true for the cognitive group, and this tendency for higher acceptances did not switch to passive waiting as predicted for the within-subjects experiment. These hypotheses originally stemmed from prevalent proposals that incentivized work should become more appealing [Eisenberger; Inzlicht], but that people would opt for the easiest available way to achieve desired outcomes [@Hull1943; Olivola]. Instead, our novel use of foraging paradigms, in addition to our inclusion of types of effort that are often studied separately, highlighted task-specific attributes contributing to the attractive or aversive nature of demands. The fact that predicted stable preferences for cognitive effort carried over onto the beginning of the within-subjects experiment further supported this idea. However, the unexpected gradual convergence in acceptances in the second experiment emphasized that these features could be contextually molded.

  These choice dynamics ruled out a number of common explanatory candidates. For instance, differences in choice stability between experiments suggest that a demand's cost is not intrinsic. The gradual choice convergence in the second experiment also discard the effects of boredom, which would have accentuated the appeal of active tasks over time, or the possibility that all our participants were high on need for cognition (i.e. Ceyda; Laura; Cacioppo et al 1996). Even though fatigue could explain the decrease in effort acceptances across experiments, we would have expected tired participants to turn to wait trials over time in order to preserve earning rates. We also ruled out the sunk cost effects used to explain overharvesting (Wikenheiser2013; Sweiss2018), as participants expressed their choices as early as possible, resulting in occassional underharvesting. Finally, the observed differences in acceptances between effort types cannot be explained by reward alone, as demands were rewarded at the same rate.

  (**hold off on this paragraph until modeling is done**) We do not discard the influence of these elements in daily effort evaluations. Instead, we computationally highlight that effort can became valuable by shortening the subjective elapsed time of trials, a perception that could be recalibrated as individuals experienced multiple forms of demands (Cognitive effort gained value by making trials feel shorter). Alterations in subjective time sensitivity have been shown to play a role in temporal discounting [@Mckerchar2003], and are often experienced when we are fully immersed in a task [@czksdmsdkjdc]. The shorter time sensitivity of cognitive effort could be due to its constant recruitment of working memory, impeding estimations of the nominal time, or because of the discrete nature of the task (versus a continuous configuration of the physical and wait tasks). These results expand our understanding of the attractiveness of effort, bringing forth outcome-independent features that can be leveraged to motivate effort engagement in diverse everyday scenarios. ....This could be because: 1) discrete nature of cog taska llowed for counting; 2) flow; 3) persistent recruitment of working memory (Buhi and Gershman). Talk about future experiments: if this is subjective time, changing the pacing of the tasks should change the acceptance rates for cognitive effort.

  Our foraging paradigms allowed us to control for common shortfalls in the existing literature. For example, even though engaging with a demanding offer encompasses evaluating, pursuing, and receiving an outcome, investigations of the pursuit and receipt stages have been limited. Many studies have sought to isolate the evaluation stage by presenting subjects with choices whose results do not come into effect until the end of the experimental session (with delays spanning up to months), deliberately setting aside the question of how the ongoing experience of cost and outcome affects subsequent choices [5,11,15,16,19]. However, discrepancies between expected and experienced value are well known [20], and choosing immediate outcomes could be due to factors other than time preference, such as the risk of being unable to redeem the future rewards, or hoping to find better offers in the future [3,21,22; hayden, mobbs?]. Here we gave participants full knowledge of the available options, and provided an opportunity to actively experience the demands they accepted. Another challenge in comparing effort and delay is that increments in the required effort are often paired with a longer execution time. This is often circumvented by comparing discounting functions and points at which individuals become indifferent to the increment in rewards []. Instead, we eliminated this possibility by fixating the handling time for all demands, which allowed us to probe people's preferences directly. In addition, the between-subjects experiment provided a baseline of foraging behavior, particularly in the traditional experimental case of foraging by waiting for the next reward []. This baseline helped us interpret the findings of the second experiment. Finally, the existence of optimal strategies let us know that people acted as if effort was attractive. This was evidenced by both the lower earnings by the cognitive effort group, and by the base OC model. Of note is that the optimality of the strategies produced by each cost depended on the richness of the block. This can help explain why the wait group earned the most despite having similar acceptance rates to the physical group, as the former was closer to optimality in the richest environment.

@@

- A potential concern is that our cognitive and physical demands were not phenomenologically effortful.
- Note that the calibration of efforts were different. We adjusted to people's physical capabilities, but trained them in the cognitive tasks until they reached criterion. 
- We still think that cognitive and physical manipulations were effortful. Our physical effort task has been repeatedly used to probe effort []. Similar task switching has been shown to feel cognitively effortful by recruiting cognitive control (Kool, some amitai). Furthermore, the effortful nature of our manipulations is evidenced by the steeper decrease in acceptances over time at longer handling times (Treadway). There are many reasons why such cognitive manipulations can be effortful, which are often centered in the amount of control. Or the OC. 
- Efficacy, along with reward amount, can influence how much control is exherted. AFAIK, Eisenberger1982 provides evidence that task diversity motivates further effort engagement.
- But here we do not focus on choosing the *amount* of cognitive control to be exherted during performance, which can be due to efficacy, agency, or reward amount. Instead, we focus on what makes people *pursue* the task in spite of how effortful it is.

@@

@@
